Evaluation of dynamic resource allocation strategies for overlay Apache Spark clusters on HPC \tristan{``Overlay'' and ``Apache Spark'' are redundant}

### 2019-05-03 paper.tex

\documentclass{IEEEtran}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\begin{document}
\title{Evaluation of dynamic resource allocation strategies for overlay Apache
       Spark clusters on HPC}
\author{
    \IEEEauthorblockN{
        Val\'erie Hayot-Sasson and Tristan Glatard
    }
    \IEEEauthorblockA{}
}
\maketitle

@@

With the increasing volume of data being collected and shared, Big Data has become
an important area of focus to many scientific domains. As is the case with neuroimaging,
dataset available publicly now reach up to Petabytes in size~\cite{UKBiobank, hcp}. Average workstations found in research lab would not be
able to support the processing of such large dataset, leaving researchers to look 
into cloud or High-Performance Computing~(HPC) solutions. Whereas both options are
suitable for the processing of scientific Big Data, HPC clusters remain the more 
sought after solution, potentially due to its cost-effectiveness.

Scientific workflows may be quite compute intensive, requiring more than a few
hours to complete (\cite{freesurfer, fmriprep, a paper which mentions duration}).
The coupling of large datasets with compute-intensive workflows may result in 
unreasonably long processing times. Such limitations restrict the amount data 
utilized in scientific experiments. To reduce the effects of Big Data on processing
times, researchers have begun to move towards employing Big Data frameworks for their
analyses. Examples of Big Data frameworks including Hadoop MapReduce~\cite{Hadoop},
Apache Spark~\cite{Spark} and Dask~\cite{Dask}. However, these frameworks were 
designed with dedicated commodity infrastructure in mind, and, with the exception
of Dask, do not support traditional HPC schedulers.

Within High-Performance Computing clusters, many users processing a variety of different
applications, share resources. These shared resources include storage (e.g. Parallel file system) and 
compute node resources (memory, cpu, local disk space, etc.). In order for users to 
gain access to shared compute resources, HPC clusters typically employ schedulers for 
batch resource provisioning. Examples of such schedulers include PBS, SGE, Slurm and 
TORQUE~\cite{schedulers}. These schedulers require that the user specify the amount
of resources required and the duration that they require the resources for. The users 
are then placed in a queue at a position based on the user's priority and resource 
availability.

There are several limitations to running applications on batch HPC schedulers, 
particularly in the face of Big Data. For instance, users may not necessarily 
know the exact requirements of the applications. With applications that are 
both data- and compute-intensive, this may result in the user overestimating the
resources required and the duration that these resources are required for. Frequent
high overestimates may create bottlenecks in the system, preventing other users
from occupying allocated idle resources. In contrast, users may underestimate required 
resources. This may subsequently result in many failures within the application leading to
additional re-executions of the same application. This is problematic for the user
as their priority may be affected by the cluster usage. Additionally, the user may 
require more time on the cluster to relaunch non-checkpointed and interrupted 
tasks. Furthermore, Big Data frameworks, such as Apache Spark, do not interact naturally
with HPC schedulers and typically rely on other schedulers for resource provisioning
and task scheduling. Therefore, to run Big Data applications on HPC schedulers, 
it is also necessary to start an overlay cluster.

In contrast to static resource provisioning, as is the case with batch HPC schedulers, 
dynamic resource provisioning, known as pilot job or glide-in scheduling, exists. 
This dynamic resource provisioning allows the user to specify a minimum and maximum 
amount of resources, and the cluster can expand and contract depending on application requirements.
Each pilot may consist of the minimal amount of resources required to launch the 
application. As each pilot would request allocation of fewer resources than that of a batch
request, Pilot Jobs would typically spend less time on the resource allocation 
queue. Furthermore, although individual pilots may expire due to walltimes, the dynamic
nature of pilots allows the entire application to surpass walltime limitations, as
additional pilots may be added to the pilot queue until application completion. 
Examples of HPC schedulers employing pilot strategies include HTCondor~\cite{htcondor}
and DIRAC~\cite{DIRAC}. \todo{maybe mention PSOM and Pegasus here}

Apache Spark is a popular Big Data framework, commonly used in both industrial
and academic settings. Although it is a Scala-based framework, it also has APIs
for Java, Python (PySpark) and R. Spark's Resilient Distributed Dataset (RDD) abstraction enabled
in-memory processing of pipelines by co-locating tasks and data, which provided important performance improvements compared to its predecessors.
 Through the use of RDDs, it also became possible to execute iterative 
workflows -- something not easily doable in older frameworks such as Hadoop MapReduce.
Schedulers for Spark include its built-in standalone schedule, Yet Another Resource Negotiator (YARN),
and Mesos~\cite{yarn, mesos}.

Projects such as Thunder~\cite{thunder}, a PySpark-compatible library 
of image processing tools, have been developed in various fields 
although they are still not widely adopted, possibly due to the 
difficulties in adapting Big Data frameworks on HPC \todo{better 
explain the difficulties...setting walltimes etc. also explain 
neuroimaging 
applications}. In addition, Spark has been used in 
various research projects to process neuroimaging data \cite{Boubela, 
ariel's paper, maybe the simulation one}. \tristan{in this paragraph you could refer
to your pre-print for more information.}

@@

Pilot scheduling techniques have already been applied to scientific workflow engines, such as
Pegasus, PSOM and RADICAL-Pilot~\cite{}. While there have been some efforts on
implementing pilot scheduling approaches for Apache Spark\cite{jha and spark on pbs paper},
research is limited and none of the them detail the effects of an overlay cluster
on overall performance and robustness.

@@

\TG{The introduction is still very much geared toward the presentation of 
the SPA software. Instead, I think it should introduce the question whether pilot jobs
are useful at all to overlay schedulers, explain why they may be, and why they may not (there is already an overlay scheduler,
and resource heterogeneity is limited). You could explain that the traditional benefits of pilot-job systems are:
\begin{itemize}
    \item Address worker node heterogeneity
    \item Address worker node failures (software, incl walltime expirations, or hardware)
    \item Address variable queuing times
    \item Tolerant to resource requirement overestimation, in particular walltime and CPU
    \item Increase workload distribution (use more nodes)
\end{itemize}
Among these, (1) and to some extent (2) are already addressed by an overlay scheduler.

>>>>

\documentclass{IEEEtran}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{ulem} % for \uwave
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}

@@

\begin{document}
\title{Evaluation of dynamic resource allocation strategies for overlay Apache
       Spark clusters on HPC \tristan{``Overlay'' and ``Apache Spark'' are redundant}}
\author{
    \IEEEauthorblockN{
        Val\'erie Hayot-Sasson and Tristan Glatard
    }
    \IEEEauthorblockA{}
}
\maketitle

@@

With the increasing volume of data being collected and shared, Big Data has become
an important area of focus to many scientific domains. As is the case with neuroimaging,
datasets available publicly now reach up to Petabytes in size~\cite{UKBiobank, hcp}. Average workstations found in research labs would not be
able to support the processing of such large datasets, leaving researchers to look 
into cloud or High-Performance Computing~(HPC) solutions. Whereas both options are
suitable for the processing of scientific Big Data, HPC clusters remain the more 
sought after solution, potentially due to their cost-effectiveness.

Scientific workflows may be quite compute intensive, requiring more than a few
hours to complete (\cite{freesurfer, fmriprep, a paper which mentions duration}).
The coupling of large datasets with compute-intensive workflows may result in 
unreasonably long processing times, which restricts the amount of data 
utilized in scientific experiments. To reduce the effects of Big Data on processing
times, researchers have begun to move towards employing Big Data frameworks for their
analyses, such as Hadoop MapReduce~\cite{Hadoop},
Apache Spark~\cite{Spark} and Dask~\cite{Dask}. However, these frameworks were 
designed with dedicated commodity infrastructure in mind, and, with the exception
of Dask, do not support traditional HPC schedulers.

Within HPC clusters, many users share resources to process a variety of different
applications. These shared resources include storage (e.g. parallel file system) and 
compute node resources (memory, CPU, local disk space, etc.). In order for users to 
gain access to shared compute resources, HPC clusters typically employ schedulers for 
batch resource provisioning. Examples of such schedulers include PBS, SGE, Slurm and 
TORQUE~\cite{schedulers}. These schedulers require that the user specify the amount
of resources required and the duration that they require the resources for. The users 
are then placed in a queue at a position based on the user's priority and resource 
availability.

There are several limitations to running applications on batch HPC schedulers, 
particularly in the face of Big Data. For instance, users may not necessarily 
know the exact requirements of the applications. With applications that are 
both data- and compute-intensive, this may result in the user overestimating the
resources required and the duration that these resources are required for. Frequent
high overestimates may create bottlenecks in the system, preventing other users
from occupying allocated idle resources. In contrast, users may underestimate required 
resources, leading the batch scheduler to kill applications due to exceeded resource quotas.
This may subsequently result in many failures within the application leading to
additional re-executions of the same application. This is problematic for the user
as their priority may be affected by the cluster usage. Additionally, the user may 
require more time on the cluster to relaunch non-checkpointed and interrupted 
tasks. \uwave{Furthermore, Big Data frameworks, such as Apache Spark, do not interact naturally
with HPC schedulers and typically rely on other schedulers for resource provisioning
and task scheduling. Therefore, to run Big Data applications on HPC schedulers, 
it is also necessary to start an overlay cluster.} \tristan{The previous 2 sentences are a bit unclear. You should explain
what an overlay cluster means here. ``Interact with other schedulers'' is vague. This point should end up explaining
what static resource provisioning means for Spark, i.e., the Compute Canada approach. It might even be a separate paragraph.
You could merge here the paragraph that currently explains that later.}

In contrast to static resource provisioning, as is the case with batch HPC schedulers, 
dynamic resource provisioning, known as pilot job or glide-in scheduling, exists. 
Dynamic resource provisioning allows the user to specify a minimum and maximum 
amount of resources, and the cluster can expand and contract depending on application requirements.
\tristan{Add one sentence to define what's a pilot}. 
Each pilot may consist of the minimal amount of resources required to launch the 
application. As each pilot would request allocation of fewer resources than that of a batch
request, Pilot Jobs would typically spend less time on the resource allocation 
queue. Furthermore, although individual pilots may expire due to walltimes, the dynamic
nature of pilots allows the entire application to surpass walltime limitations, as
additional pilots may be added to the pilot queue until application completion. 
Examples of HPC schedulers employing pilot strategies include HTCondor~\cite{htcondor}
and DIRAC~\cite{DIRAC}. \todo{maybe mention PSOM and Pegasus here \tristan{Yes for PSOM as it's neuroimaging. Pegasus
only uses pilots through condor afaik.}}

\tristan{I think this could be the 3rd paragraph.} Apache Spark is a popular Big Data framework, commonly used in both industrial
and academic settings. Although it is a Scala-based framework, it also has APIs
for Java, Python (PySpark) and R. Spark's Resilient Distributed Dataset (RDD) abstraction enabled
in-memory processing of pipelines by co-locating tasks and data, which provided important performance improvements compared to its predecessors.
 Through the use of RDDs, it also became possible to execute iterative 
workflows -- something not easily doable in older frameworks such as Hadoop MapReduce.
Schedulers for Spark include its built-in standalone schedule, Yet Another Resource Negotiator (YARN),
and Mesos~\cite{yarn, mesos}. \tristan{You could cite Ariel's and your paper to say that Spark is sometimes useful in 
neuroimaging on a cluster.}

Projects such as Thunder~\cite{thunder}, a PySpark-compatible library 
of image processing tools, have been developed in various fields 
although they are still not widely adopted, possibly due to the 
difficulties in adapting Big Data frameworks on HPC \todo{better 
explain the difficulties...setting walltimes etc. also explain 
neuroimaging 
applications}. In addition, Spark has been used in 
various research projects to process neuroimaging data \cite{Boubela, 
ariel's paper, maybe the simulation one}. \tristan{in this paragraph you could refer
to your pre-print for more information.} \tristan{I would summarize that in 1 or 2 sentences at the end of the previous
paragraph.}

@@

Pilot scheduling techniques have already been applied to scientific workflow engines, such as
Pegasus, PSOM and RADICAL-Pilot~\cite{}. While there have been some efforts on
implementing pilot scheduling approaches for Apache Spark\cite{jha and spark on pbs paper},
research is limited and none of the them detail the effects of an overlay cluster
on overall performance and robustness. 
\tristan{This paragraph should write the scientific question of your paper more clearly: is it useful to 
use a pilot job framework to run Spark workflows on HPC? Also, this comes a bit late in the intro, it would be good
to precise the question early on, and the explain it in length as you do.}

@@

\tristan{Your text now mentions points 3 and 4 below, but doesn't talk about 1 and 2 and the fact that
an overlay cluster already deals with that. Something that wasn't mentioned in my initial list was that pilot jobs
also reduce the number of tasks seen by the cluster (this is also addressed by an overlay scheduler). It would be interesting
to explain that many of the interests of pilot jobs are already addressed by overlay scheduling, hence the question whether
pilot jobs are useful at all in this context.}
\TG{The introduction is still very much geared toward the presentation of 
the SPA software. Instead, I think it should introduce the question whether pilot jobs
are useful at all to overlay schedulers, explain why they may be, and why they may not (there is already an overlay scheduler,
and resource heterogeneity is limited). You could explain that the traditional benefits of pilot-job systems are:
\begin{itemize}
    \item Address worker node heterogeneity
    \item Address worker node failures (software, incl walltime expirations, or hardware)
    \item Address variable queuing times
    \item Tolerant to resource requirement overestimation, in particular walltime and CPU
    \item Increase workload distribution (use more nodes)
\end{itemize}
Among these, (1) and to some extent (2) are already addressed by an overlay scheduler.
